---
title: "10_Tidy_Data"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice tidying data

### Find mean

```{r}
library(tidyverse)
library(babynames)
mean(table1$cases)
```

### Compute ratios

```{r}
table1$cases / table1$population
```

### Use gather to convert wide data to long data

```{r}
table4a

table4a %>% gather(key = "year", value = "cases", 2, 3)
```

### Can convert using a variety of methods

```{r}
table4a %>% gather(key = "year", value = "cases", `1999`, `2000`)
table4a %>% gather(key = "year", value = "cases", -country)
table4a %>% gather(key = "year", value = "cases", one_of(c("1999", "2000")))
```

### Need to ask R to convert new column to appropriate data type by adding convert

```{r}
table4b %>% gather(key = "year", value = "population", -country, convert = TRUE)
```

### Practice using gather on larger dataset

```{r}
cases2 <- read.csv("/Users/ashafil/Documents/R/cases2.csv")

cases2

cases2 %>% gather(key = "year", value = "cases", 4, 5, 6)
```

### Can use 'spread' and change what columns used askey and what column used as values

```{r}
pollution <-read.csv("/Users/ashafil/Documents/R/pollution.csv")

pollution

pollution %>% spread(key = size, value = amount)

table2

table2 %>% spread(key = type, value = count)
```

### Note: gather uses quotations around the  key and value arguments whereas spread does not, gather refers to columns that do not exist while spread points to columns that do exist

### Review making a plot to compare boy and girl babynames over time

```{r}
babynames %>%
  group_by(year, sex) %>% 
  summarise(total = sum(n)) %>% 
  ggplot() +
  geom_line(aes(year, total, color = sex))
```

### Better way to look at it is through using a ratio

```{r}
babynames %>%
  group_by(year, sex) %>% 
  summarise(total = sum(n))
```

### But easier to calculate if reshape data

```{r}
babynames %>%
  group_by(year, sex) %>% 
  summarise(total = sum(n)) %>%
  spread(sex, total) %>%
  mutate(ratio = M / F) %>%
  ggplot(aes(year, ratio)) + 
  geom_line()
```

