---
title: "11_Separate_Unite_Columns"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice separating and uniting columns
### Import data

```{r}
library(dplyr)
library(tidyverse)
hurricanes <- read.csv("hurricanes.csv")

hurricanes
```

### What if want to separate the elements of the date?

```{r}
hurricanes %>% 
  separate(col = date, into = c("year","month","day"), sep = "-")
```

### Can further separate by location defined by integers

```{r}
hurricanes %>% 
  separate(col = date, into = c("year","month","day")) %>% 
  separate(col = year, into = c("century", "year"), sep = 2)
```

### Also can separate using 'convert = TRUE'

```{r}
hurricanes %>% 
  separate(col = date, into = c("year","month","day"), convert = TRUE)
```

### Can retain original column and have new column

```{r}
hurricanes %>% 
  separate(col = date, into = c("year","month","day"), convert = TRUE, remove = FALSE)
```

### Can put columns together by using 'unite'

```{r}
hurricanes %>%
  separate(date, c("year", "month", "day"), sep = "-") %>%
  unite(col = "date", month, day, year, sep = ":")

hurricanes %>%
  separate(date, c("year", "month", "day"), sep = "-") %>%
  unite(col = "date", month, day, year, sep = "/")
```

### Can filter if just want data from the 19th century

```{r}
hurricanes %>%
  separate(col = date, c("century", "rest"), sep = 2) %>%
  filter(century == 19) %>% 
  unite(col = "date", century, rest, sep = "")
```

### When working in long format, if there is N/A, you may be able to remove the whole column since it does not have any data in it

```{r}
data %>% drop_na(column1, column2)
```

