---
title: "12_Join_Datasets"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice joining datasets in different ways

### Compare average flight delay time by airline

```{r}
library("nycflights13")
library(dplyr)
library(tidyr)

flights

flights %>% 
  drop_na(arr_delay) %>% 
  group_by(carrier) %>% 
  summarise(avg_delay = mean(arr_delay)) %>% 
  arrange(desc(avg_delay))
```

### Want to join this outcome with the airlines dataset to know the airline name and not just the code 

```{r}
flights %>% 
  drop_na(arr_delay) %>% 
  group_by(carrier) %>% 
  summarise(avg_delay = mean(arr_delay)) %>% 
  arrange(desc(avg_delay)) %>% 
  left_join(airlines, by = "carrier") %>% 
  select(name, avg_delay)
```

### See which destinations have largest average arrival delays

```{r}
flights %>% 
  drop_na(arr_delay) %>%
  group_by(dest) %>%
  summarise(avg_delay = mean(arr_delay)) %>%
  arrange(desc(avg_delay))
```

### Want to combine datasets but they have different column names (airports uses faa instead of dest)

```{r}
flights %>% 
  drop_na(arr_delay) %>%
  group_by(dest) %>%
  summarise(avg_delay = mean(arr_delay)) %>%
  arrange(desc(avg_delay)) %>% 
  left_join(airports, by = c("dest" = "faa")) %>% 
  select(name, avg_delay)
```

### Semi-join returns every row of first data set that matches second dataset but does not add anything from the second dataset

```{r}
band <- read.csv("band.csv")
instrument <- read.csv("instrument.csv")

band %>% semi_join(instrument, by = "name")
```

### Anti-join does opposite, returns rows in first dataset that don't match in the second data set

```{r}
band %>% anti_join(instrument, by = "name")
```

### Distinct gives you values that are unique in a column

```{r}
instrument %>% distinct(plays)
```

### How many airports does New York City connect to?

```{r}
flights %>% 
  distinct(dest)

flights %>% 
  distinct(dest) %>% 
  left_join(airports, by = c("dest" = "faa")) %>% 
  select(name)
```

### Some codes did not have a match with the airports, so we use anti_join to figure out which ones they are

```{r}
flights %>%
  distinct(dest) %>% 
  anti_join(airports, by = c("dest" = "faa"))
```

### Let's find all of the codes that do match an airport

```{r}
flights %>% 
  semi_join(airports, by = c("dest" = "faa")) 
```

### What if want flights that depart in January on JetBlue or depoart February on Southwest?

```{r}
criteria <- tribble(
  ~month, ~carrier,
  1,     "B6", # B6 = JetBlue
  2,     "WN"  # WN = Southwest
)

criteria

flights %>% 
  semi_join(criteria, by = c("month", "carrier"))
```

### If data contains same observations in same order, you can combine them with 'bind_cols'

```{r}
beatles1 <- read.csv("beatles1.csv")
beatles2 <- read.csv("beatles2.csv")

beatles1 %>% bind_cols(beatles2)
```

### Or with 'bind_rows'

```{r}
band2 <- read.csv("band2.csv")

band %>% bind_rows(band2)
```

#When combining rows, it's useful to add new column showing where each row came from 

```{r}
bands <- list(df1 = band, 
              df2 = band2)

bands %>% bind_rows(.id = "origin")
```

### Union returns every row in data sets but removes duplicates

```{r}
band %>% union(band2)
```

### Intersect returns that only appear in both (also removes duplicates)

```{r}
band %>% intersect(band2)
```

### Setdiff returns all of the rows that appear in the first data set but not the second.

```{r}
band %>% setdiff(band2)

band2 %>% setdiff(band)
```

