---
title: "13_Iterate"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice methods to iterate vectors or lists

### Nonrecursive vectors- each vector is single value and the same type

```{r}
numbers <- c(1, 2, 3, 4, 5)
log(numbers)
```

### Recursive vectors (aka lists) - each element can be anything, even other vectors

```{r}
list1 <- list(1, 2, 3, 4, 5)
chars <- c("a", "b", "c")
list2 <- list(1, chars)
list2
```

### Can extract each element of list with [[i]]

```{r}
list2[[2]]
```

### Using purr package from tidyverse 

```{r}
library(tidyverse)
```

### Map function to applies R function to each element of R vector
### Example: find absolute volume

```{r}
negatives <- list(-25, -31, -111)
map(negatives, abs)

list1 <- list(1, 2, 3, 4, 5)
map(list1,log)
```

### Can use pipes

```{r}
list1 %>% 
 map(log) %>% 
  map(round)

list1 %>% 
 map(log) %>% 
 map(round, digits = 2)

map(list1,log, base = 2)
```

### Want to return results as vector and not list, can use map_dbl

```{r}
map_dbl(list1, log)

```

### Map have multiple versions: character vector (map_chr), double/numeric (map_dbl), dataframe-column binded (map_dfc),dataframe-row binded (map_dfr), integer (map_int), logical (map_lgl), functions that 
### plot graphs or saves files (walk)

```{r}
strings <- list("Mary", "John", "Jill")
```

### Note str_sub makes all arguments the same length 

```{r}
map_chr(strings, str_sub, start = 1, end = 1)
```

### Pluck extracts element from a list by name or position and returns as vector

```{r}
list1 <- list(
  numbers = 1:3,
  letters = c("a", "b", "c"),
  logicals = c(TRUE, FALSE)
)

pluck(list1, 1) # list1 %>% pluck(1)
pluck(list1, "numbers") # list1 %>% pluck("numbers")
```

### Making a data frame

```{r}
df <- data.frame(
  name = c("John", "Mary", "Mike"),
  age = c(20, 21, 22),
  height = c(69, 64, 72)
)

df
```

### Another way to add column

```{r}
df$height2 <- list(
  c(feet = 5, inches = 9),
  c(feet = 5, inches = 4),
  c(feet = 6, inches = 0)
)

df
```

