---
title: "2_Arrange_Select_Filter"
author: "Joanne Fil"
date: "7/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice using 'arrange', 'select' and 'filter' to manipulate the dataset

### Install packages and view the data

```{r babynames}
library(babynames)
library(tibble)
library(tidyverse)
head(babynames)
```

#### How to select various columns:
### Select columns except n

```{r}
select(babynames, -n)
```

### Select between certain columns (inclusive)

```{r}
select(babynames, year:name)
```

### Select columns that contains a specific string

```{r}
select(babynames, contains("se"))
```

### Select columns that end with a specific string

```{r}
select(babynames, ends_with("r"))
```

### How to filter names:

```{r}
filter(babynames, name == "Elizabeth")
```

### Filter name and proportion

```{r}
filter(babynames, name == "Elizabeth" & prop > 0.001)
```

### Filter name or proportion

```{r}
filter(babynames, name == "Elizabeth" | prop > 0.001)
```

### Filter year

```{r}
filter(babynames, year == 1880 | year == 1900 | year == 1992)
filter(babynames, year %in% c(1880, 1900, 1992))
```

### Arrange order

```{r}
arrange(babynames, name)
arrange(babynames, year, desc(prop))
arrange(babynames,desc(prop))
```

### Filter, select, and arrange

```{r}
boys_2017 <- filter(babynames, year == 2017, sex == "M")
head(boys_2017)
boys_2017 <- select(boys_2017, name, n)
head(boys_2017)
boys_2017 <- arrange(boys_2017, desc(n))
head(boys_2017)
```

### Faster way to run codes is by using piper operator

```{r}
babynames %>% 
  filter(name == "Elizabeth", sex == "F") %>%
  summarize(total = sum(n), max = max(n), min = min(n), mean = mean(n))
```

### Practice: filter specific name and plot it

```{r}
Eliz <- filter(babynames, name == "Elizabeth") 

ggplot(data = Eliz) +
  geom_line(mapping = aes(x = year, y = n))
```













