---
title: "3_Bar_Chart"
author: "Joanne Fil"
date: "7/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice creating different types of bar charts

### View the data

```{r}
library(dplyr)
library(ggplot2)
head(diamonds)
diamonds %>% count(cut)
```

### Map to the number in each category of 'cut'

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

### Change color of of each cut type

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

### Change width

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut), width = 0.3)
```

### To make columns without just counting, use 'geom_col'

```{r}
head(pressure)
ggplot(data = pressure) +
  geom_col(mapping = aes(x = temperature, y = pressure))
```

### Stacked bar chart

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))

```

### Stacked bar chart where it takes up the entire y axis

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

### Stacked bar chart without space in between columns

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill", width = 1)
```

### Use dodge to place rectangles that have the same cut value beside each other

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

### Facet - split into subplots; vertical by first value and horizontal by second value

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(clarity ~ cut)

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(. ~ cut)

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(clarity ~ .)
```

### Facet_wrap - facet a plot on a single variable

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut)) +
  facet_wrap( ~ cut)
```

### Facet_wrap can let range of x, y, or both vary

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut)) +
  facet_wrap( ~ cut, scales = "free_x")

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut)) +
  facet_wrap( ~ cut, scales = "free_y")

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut)) +
  facet_wrap( ~ cut, scales = "free")
```


