---
title: "4_Histogram"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice creating different types of histograms

### Making a histogram

```{r}
library(ggplot2)
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat))
```

### Can change bin width

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)
```

### Can change number of bins

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), bins = 50)
```

### Can move the bins left and right along the x axis

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), bins = 10, boundary = 0)
```

### Also can stack

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, fill = cut), binwidth = 1000, boundary = 0)
```

### Frequency polygon, uses a line to display the same information as a histogram

```{r}
ggplot(data = diamonds) +
  geom_freqpoly(mapping = aes(x = carat), binwidth = 0.25, boundary = 0)
```

### Split the frequency polygon to multiple lines

```{r}
ggplot(data = diamonds) +
  geom_freqpoly(mapping = aes(x = price, color = cut), binwidth = 1000, boundary = 0)
```

### Density curve

```{r}
ggplot(data = diamonds) +
  geom_density(mapping = aes(x = price, color = cut))
```

### Dotplot represents each observation with a dot and then stacks dots within bins to create the semblance of a histogram.

```{r}
ggplot(data = mpg) +
  geom_dotplot(mapping = aes(x = displ), dotsize = 0.5, stackdir = "up", stackratio = 1.1)
```

### Can facet histograms

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, fill = cut), binwidth = 1000, boundary = 0) +
  facet_wrap( ~cut)
```

