---
title: "5_Boxplot_Counts_Heatmap"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practce using boxplots, counts, and heatmaps

### Boxplot

```{r}
library(ggplot2)
library(dplyr)
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy))
```

### Boxplot with diamond data - many outliers

```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = price))
```

### Can change how outliers look

```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = price), outlier.shape = 24, 
               outlier.fill = "white", outlier.stroke = 0.25)
```

### Can make outliers disappear

```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = price), outlier.alpha = 0)
```

### Can split by group with vertical boxplots

```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = carat, y = price, 
                             group = cut_width(carat, width = 0.25)))
```

### Can split by group with horizontal boxplots

```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy)) + 
  coord_flip()
```

### Distribute data with dots

```{r}
ggplot(data = mpg) +
  geom_dotplot(mapping = aes(x = class, y = hwy), binaxis = "y", 
               dotsize = 0.5, binwidth = 1)
```

```{r}
ggplot(data = mpg) +
  geom_dotplot(mapping = aes(x = class, y = hwy), binaxis = "y", 
               dotsize = 0.5, binwidth = 1, stackdir = 'center')
```

### Violin plot to use densities to draw a smoothed version of the dotplot

```{r}
ggplot(data = mpg) +
  geom_violin(mapping = aes(x = class, y = hwy))
```

### Violin plot with 25th, 50th, and 75th percentile

```{r}
ggplot(data = mpg) +
  geom_violin(mapping = aes(x = class, y = hwy),
              draw_quantiles = c(0.25, 0.5, 0.75))
```

### Interaction of continuous variable and two categorical variables

```{r}
ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = clarity))

```

### Count - compute count values

```{r}
diamonds %>% 
  count(cut, clarity)
```

### Heatmap

```{r}
diamonds %>% 
  count(color, cut) %>% 
  ggplot(mapping = aes(x = color, y = cut)) +
  geom_tile(mapping = aes(fill = n))
```



