---
title: "6_Scatterplot"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice using scatterplots

### Scatterplot

```{r}
library(ggplot2)
library(dplyr)
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
  geom_point(mapping = aes(x = mean_cty, y = mean_hwy))
```

### Scatterplot with words, either geom_text or geom_label

```{r}
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
  geom_text(mapping = aes(x = mean_cty, y = mean_hwy, label = class))

mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
  geom_label(mapping = aes(x = mean_cty, y = mean_hwy, label = class))
```

### Use 'model' to fit linear (lm) or generalized liner (glm) models

```{r}
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
  geom_smooth(mapping = aes(x = mean_cty, y = mean_hwy), method = lm)
```

### Scatterplot with raw data and a trend line

```{r}
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
  geom_point(mapping = aes(x = mean_cty, y = mean_hwy)) +
  geom_smooth(mapping = aes(x = mean_cty, y = mean_hwy), method = lm) 
```

### Arrange labels within a plot

```{r}
library("ggrepel")
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
  geom_point(mapping = aes(x = mean_cty, y = mean_hwy)) +
  geom_smooth(mapping = aes(x = mean_cty, y = mean_hwy), method = lm) +
  geom_label_repel(mapping = aes(x = mean_cty, y = mean_hwy, label = class))
```

### More efficient way to write above code

```{r}
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot(mapping = aes(x = mean_cty, y = mean_hwy)) +
  geom_point() +
  geom_smooth(method = lm) +
  geom_label_repel(mapping = aes(label = class))
```

### Dots and lines

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(mapping = aes(color = class), se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth()
```

### Highlight certain points in color

```{r}
mpg2 <- filter(mpg, class == "2seater")

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_point(data = mpg2, color = "red", size = 2)
```

### Highlight certain points using labels

```{r}
mpg3 <- filter(mpg, hwy > 40)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_label_repel(data = mpg3, mapping = aes(label = class))
```

### Can pull up last plot and add onto it

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point()

last_plot() +
  geom_smooth()

last_plot() +
  geom_smooth(method = lm, color = "purple")
```

### Display one dimensional marginal distribution of each variable as collection of tickmarks

```{r}
ggplot(data = faithful, mapping = aes(x = waiting, y = eruptions)) +
  geom_point() +
  geom_rug()
```

### Jitter to add small amount of noise - inspect patterns in heavily overlapping data

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = hwy))

ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = class, y = hwy))
```

### Boxplot and jitter

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(width = 0)

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(width = 0) +
  coord_flip()
```

### Can flip to use polar coordinates

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut), width = 1) 

last_plot() +
  coord_polar()
```

### Log-log plot

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = log(carat), y = log(price))) 
```

### Log axis

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price)) +
  coord_trans(x = "log", y = "log")
```

