---
title: "7_Line_Graph"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice using line graphs

### Line graph 

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
ggplot(economics) +
  geom_line(mapping = aes(x = date, y = unemploy))
```

### Line graph but fills area under

```{r}
ggplot(economics) +
  geom_area(mapping = aes(x = date, y = unemploy))
```

### Line graph but fills area under with color

```{r}
ggplot(economics) +
  geom_area(mapping = aes(x = date, y = unemploy), fill = "blue")
```

```{r}
library("gapminder")

asia <- read.csv("asia.csv", header = TRUE)
```

### Whipsaw

```{r}
ggplot(asia) +
  geom_line(mapping = aes(x = year, y = gdpPercap))

```

### Scatterplot lines

```{r}
ggplot(asia) +
  geom_point(mapping = aes(x = year, y = gdpPercap))
```

### Separate by line

```{r}
ggplot(asia) +
  geom_line(mapping = aes(x = year, y = gdpPercap, group = country))
```

### Separate by line and color 

```{r}
ggplot(asia) +
  geom_line(mapping = aes(x = year, y = gdpPercap, color = country))
```

### Separate by different line types and color 

```{r}
ggplot(asia) +
  geom_line(mapping = aes(x = year, y = lifeExp, color = country, linetype = country))

```

### Overlap of filled lines 

```{r}
ggplot(asia) +
  geom_area(mapping = aes(x = year, y = lifeExp, fill = country), position = "identity", alpha = 0.3)
```

### Longitude and Latitude data

```{r}
library(maps)
usa <- map_data("usa")
```

### Shows map outline

```{r}
ggplot(usa) +
  geom_path(mapping = aes(x = long, y = lat))
```

### Shows colored map

```{r}
ggplot(usa) +
  geom_polygon(mapping = aes(x = long, y = lat))

```

### Get subset of region

```{r}
fl <- map_data("state", region = "florida")
ggplot(fl) +
  geom_polygon(mapping = aes(x = long, y = lat))
```

### This one gives us strange image because contains multiple polygons and is connecting them in weird ways

```{r}
us <- map_data("state")
ggplot(us) +
  geom_polygon(mapping = aes(x = long, y = lat))
```

### This fixes it by grouping them

```{r}
ggplot(us) +
  geom_polygon(mapping = aes(x = long, y = lat, group = group))
```

### Chloropleth map

```{r}
USArrests2 <- USArrests %>% 
  rownames_to_column("region") %>% 
  mutate(region = tolower(region))

USArrests2

ggplot(USArrests2) +
  geom_map(aes(map_id = region, fill = UrbanPop), map = us) +
  expand_limits(x = us$long, y = us$lat)
```

### Fix map distortion by using coord_map()

```{r}
ggplot(USArrests2) +
  geom_map(aes(map_id = region, fill = UrbanPop), map = us) +
  expand_limits(x = us$long, y = us$lat) +
  coord_map()
```

### Can change to different projection instead of Mercator default

```{r}
ggplot(USArrests2) +
  geom_map(aes(map_id = region, fill = UrbanPop), map = us) +
  expand_limits(x = us$long, y = us$lat) +
  coord_map(projection = "sinusoidal")
```

