---
title: "9_Customized_plots"
author: "Joanne Fil"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practice with customizing plots

### Make a boxplot

```{r}
library(ggplot2)
library(dplyr)
ggplot(diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = price))

p <- ggplot(diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = price))
```

### Worst cut diamond is the lowest price, check with summarizing

```{r}
diamonds %>% 
  group_by(cut) %>% 
  summarise(median = median(price))
```

### Can zoom the y axis from 0 to 7,500

```{r}
p + ylim(0, 7500)
```

### Caution: zooming is bad for boxplots since it changes the information because it throws out some data before drawing the boxplot

### To zoom without clipping change the coordinate

```{r}
p + coord_flip(ylim = c(0, 7500))
```

### If plot doesn't call the coord_ function then plot is using Cartesian coordinates

```{r}
p + coord_cartesian(ylim = c(0, 7500))
```

### Can update by overwriting

```{r}
p <- p + coord_cartesian(ylim = c(0, 7500))
p
```

### Add title

```{r}
p + labs(title = "The title appears here")
```

### Add subtitle

```{r}
p + labs(title = "The title appears here",
         subtitle = "The subtitle appears here, slightly smaller")
```

### Add caption

```{r}
p + labs(title = "The title appears here",
         subtitle = "The subtitle appears here, slightly smaller",
         caption = "Captions appear at the bottom.")

p + labs(title = "Diamond prices by cut",
         subtitle = "Fair cut diamonds fetch the highest median price. Why?",
         caption = "Data collected by Hadley Wickham")
```

### Plot displays relationship between carat size, price, and cut

```{r}
p1 <- ggplot(data = diamonds, mapping = aes(x = carat, y = price)) +
  geom_smooth(mapping = aes(color = cut), se = FALSE) + 
  labs(title = "Carat size vs. Price",
       subtitle = "Fair cut diamonds tend to be large, but they fetch the lowest prices for most carat sizes.",
       caption = "Data by Hadley Wickham")
```

### Annotate adds a single geom to plot

```{r}
p1 + annotate("text", x = 4, y = 7500, label = "There are no cheap,\nlarge diamonds")
```

### Can use different themes

```{r}
p1 + theme_bw()
p1 + theme_classic()
p1 + theme_dark()
p1 + theme_gray()
p1 + theme_light()
p1 + theme_linedraw()
p1 + theme_minimal()
p1 + theme_void()

library("ggthemes")
p1 + theme_base()
p1 + theme_calc()
p1 + theme_economist()
p1 + theme_economist_white()
p1 + theme_excel()
p1 + theme_few()
p1 + theme_fivethirtyeight()
p1 + theme_foundation()
p1 + theme_gdocs()
p1 + theme_hc()
p1 + theme_igray()
p1 + theme_map()
p1 + theme_pander()
p1 + theme_par()
p1 + theme_solarized()
p1 + theme_solarized_2()
p1 + theme_solid()
p1 + theme_stata()
p1 + theme_tufte()
p1 + theme_wsj()
```

### Use scales, like greyscale

```{r}
p1 + scale_color_grey()
```

### Logscale

```{r}
p1 +
  scale_x_log10() + 
  scale_y_log10()
```

### Good color for color palettes for discrete variables

```{r}
p1 + scale_color_brewer()
```

### Show different palettes

```{r}
RColorBrewer::display.brewer.all()
```

### Change the color palette of the lines

```{r}
p1 + scale_color_brewer(palette = "Purples")

p1 + scale_color_brewer(palette = "Spectral")
```

### What about continuous variables?
### Use scale_color_distiller with same palettes as before

```{r}
p_cont <- ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, y = hwy, color = hwy)) +
  theme_bw()

p_cont

p_cont + scale_color_distiller(palette = "Spectral")

p_cont + scale_color_distiller(palette = "BrBG")
```

### Can also use viridis package for continuous variables and are color blind and black and white printer friendly

```{r}
library("viridis")

p_cont + scale_color_viridis()
```

### Viridis color magma

```{r}
p_cont + scale_color_viridis(option = "A")
```

### Viridis color plasma

```{r}
p_cont + scale_color_viridis(option = "B")
```

### Viridis color inferno

```{r}
p_cont + scale_color_viridis(option = "C")
```

### Viridis color magma

```{r}
p_cont + scale_color_viridis(option = "D")
```

### Change position of legend

```{r}
p_cont + theme(legend.position = "bottom")

p_cont + theme(legend.position = "top")

p_cont + theme(legend.position = "left")

p_cont + theme(legend.position = "right")
```

### Different way to organize the plot

```{r}
ggplot(data = mpg) + 
  geom_jitter(mapping = aes(x = displ, y = hwy, color = hwy)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

### Can change the labels of the legend

```{r}
p1 + scale_color_brewer(name = "Cut Grade", 
                        labels = c("Very Bad", "Bad", "Mediocre", "Nice", "Very Nice"))
```

### If need to identify default scales and then manually add that scale to the plot

```{r}
p1 + scale_color_discrete(name = "Cut Grade", labels = c("Very Bad", "Bad", "Mediocre", "Nice", "Very Nice"))

p_cont + scale_color_continuous(name = "Highway MPG") + theme(legend.position = "top")
```

### Axis labels

```{r}
p1 + scale_x_continuous(name = "Carat Size", labels = c("Zero", "One", "Two", "Three", "Four", "Five"))
```

### Example plots

```{r}
ggplot(data = diamonds, mapping = aes(x = carat, y = price)) +
  geom_point() + 
  geom_smooth(mapping = aes(color = cut), se = FALSE) +  
  labs(title = "Ideal cut diamonds command the best price for every carat size",
       subtitle = "Lines show GAM estimate of mean values for each level of cut",
       caption = "Data provided by Hadley Wickham",
       x = "Log Carat Size",
       y = "Log Price Size",
       color = "Cut Rating") +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_brewer(palette = "Greens") +
  theme_light()
```

